Containers and R: the Rockerverse and beyond

### 2019-10-21 manuscript.Rmd
Added Marwick paper citation

To set up Binder for a project, a user typically starts at an instance of a Binderhub (examples include mybinder.org and binder.pangeo.io) and passes the location of a hosted Git repository. Using internal tools such as [repo2docker](https://repo2docker.readthedocs.io/en/latest/config_files.html), the binderhub builds a Dockerfile by parsing the contents of all code contained within. While this approach works well for most run of the mill Python projects, it is not so seamless for R projects. For any R projects that use the Tidyverse suite [citation in review], the time and resources required to build all dependencies from source can often time out before completion, making it frustrating for the average R user.

[`holepunch`](https://github.com/karthik/holepunch) is a R package that was designed to remove some of these limitations and make Binder more accessible to novice R users. `holepunch` achieves this by leveraging Rocker images that contains the Tidyverse along special Jupyter dependencies, and only installs additional packages from CRAN and Bioconductor that are not already part of these images. Starting with the Binder/Tidyverse base images eliminates a large part of the build time and in most cases results in a binder instance launching within a minute. `holepunch` as a side effect also creates a DESCRIPTION file which then turns any project into a research compendium [Marwick et al]. The Dockerfile included with the project can also be used to launch a RStudio server locally independent of binder which is especially useful when more computational resources are required.

>>>>

To set up Binder for a project, a user typically starts at an instance of a Binderhub (examples include mybinder.org and binder.pangeo.io) and passes the location of a hosted Git repository. Using internal tools such as [repo2docker](https://repo2docker.readthedocs.io/en/latest/config_files.html), the binderhub builds a Dockerfile by parsing the contents of all code contained within. While this approach works well for most run of the mill Python projects, it is not so seamless for R projects. For any R projects that use the Tidyverse suite [citation in review. Will add by the time this gets submitted], the time and resources required to build all dependencies from source can often time out before completion, making it frustrating for the average R user.

[`holepunch`](https://github.com/karthik/holepunch) is a R package that was designed to remove some of these limitations and make Binder more accessible to novice R users. `holepunch` achieves this by leveraging Rocker images that contains the Tidyverse along special Jupyter dependencies, and only installs additional packages from CRAN and Bioconductor that are not already part of these images. Starting with the Binder/Tidyverse base images eliminates a large part of the build time and in most cases results in a binder instance launching within a minute. `holepunch` as a side effect also creates a DESCRIPTION file which then turns any project into a research compendium [doi:10.1080/00031305.2017.1375986]. The Dockerfile included with the project can also be used to launch a RStudio server locally independent of binder which is especially useful when more computational resources are required.

### 2019-10-21 manuscript.tex
Added Marwick paper citation

\begin{itemize}
\tightlist
\item
  Binder

  \begin{itemize}
  \tightlist
  \item
    \href{https://github.com/karthik/holepunch}{\texttt{holepunch}}
    streamlines making an R research compendium Binder-ready
    {[}@karthik{]}
  \item
    \href{https://repo2docker.readthedocs.io/en/latest/config_files.html}{\texttt{repo2docker}}
    (\texttt{install.R}, \texttt{DESCRPTION}, \texttt{runtime.txt})
    based on the Rocker images {[}@nuest{]}
  \end{itemize}
\item
  GPU {[}@noamross, @cboettig{]}
\item
  Gigantum stack
\item
  in education (ready to use teaching/learning environments)

>>>>

\begin{itemize}
\tightlist
\item
  Binder
\end{itemize}

The \href{https://mybinder.readthedocs.io/en/latest/}{Binder project},
maintained by the team behind Jupyter, makes it possible for users to
create and share their computing environments with others. A deployed
version of this service called binder hub allows anyone with access to a
web browser and an internet connection to launch a temporary instance of
these custom environments and execute any notebooks contained within.
From a reproducibility standpoint, Binder makes it exceedingly easy to
compile a paper, visualize data, and run small examples from papers or
tutorials without the need for any local installation.

To set up Binder for a project, a user typically starts at an instance
of a Binderhub (examples include mybinder.org and binder.pangeo.io) and
passes the location of a hosted Git repository. Using internal tools
such as
\href{https://repo2docker.readthedocs.io/en/latest/config_files.html}{repo2docker},
the binderhub builds a Dockerfile by parsing the contents of all code
contained within. While this approach works well for most run of the
mill Python projects, it is not so seamless for R projects. For any R
projects that use the Tidyverse suite {[}citation in review{]}, the time
and resources required to build all dependencies from source can often
time out before completion, making it frustrating for the average R
user.

\href{https://github.com/karthik/holepunch}{\texttt{holepunch}} is a R
package that was designed to remove some of these limitations and make
Binder more accessible to novice R users. \texttt{holepunch} achieves
this by leveraging Rocker images that contains the Tidyverse along
special Jupyter dependencies, and only installs additional packages from
CRAN and Bioconductor that are not already part of these images.
Starting with the Binder/Tidyverse base images eliminates a large part
of the build time and in most cases results in a binder instance
launching within a minute. \texttt{holepunch} as a side effect also
creates a DESCRIPTION file which then turns any project into a research
compendium {[}\url{doi:10.1080/00031305.2017.1375986}{]}. The Dockerfile
included with the project can also be used to launch a RStudio server
locally independent of binder which is especially useful when more
computational resources are required.

\begin{itemize}
\tightlist
\item
  \href{https://repo2docker.readthedocs.io/en/latest/config_files.html}{\texttt{repo2docker}}
  (\texttt{install.R}, \texttt{DESCRPTION}, \texttt{runtime.txt}) based
  on the Rocker images {[}@nuest{]}
\item
  GPU {[}@noamross, @cboettig{]}
\item
  Gigantum stack
\item
  in education (ready to use teaching/learning environments)
